---
title: "Regression_lab1"
author: "2229027 Jimin Lee"
output:
  word_document: default
  pdf_document: default
---
The dataset record.txt contains running records obtained from athletes from different countries in various types of athletics events (sprints and middle-distance). We have data about 55 countries (observations) and 6 records (variables): 100 meters, 200 meters, 400 meters, 800 meters, 1500 meters and 3000 meters.

### Load the dataset record.txt in R, using the function read.table

```{r}
# prepare data
setwd('/Users/jimin/Desktop/지민/ewha/2023-2/Regression/')
record <- read.table("record.txt" ,head=TRUE)
head(record, n=10)
```
### Produce summaries of the variable m800, including 
#### Numerical summaries: average, standard deviation, median and quartiles, maximum and minimum, interquartile difference
```{r}
# numerical summaries
print(summary(record$m800))
cat("sd : ", sd(record$m800), "\n")
cat("IQR : ", IQR(record$m800))
```
#### Graphical summaries: histogram and boxplot
#### What can you observe about the variable distribution?

The distribution of m800 record is right-skewed.

```{r}
# graphical summaries
hist(x=record$m800, main='Histogram of m800 record')
help(hist)
```

```{r}
boxplot(record$m800, main='boxplot of m800 record')
```
### Produce scatter plot between all the variables(m100,m200,m400,m800,m1500,m3000). What can you observe from the scatter plot? Are they correlated?
When x increases, also y increase. 
positive correlation.

```{r}
pairs(record, main = "athletes per records") # print all scatters between cols
help(pairs)
```
```{r}
cor(record) # search the correlation between x,y
```


